Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-05-12 Paper_draft/main.tex
modified:   iopart-num.bib 	modified:   main.tex

\acrodef{GW}[GW]{gravitational wave}
\acrodef{GAN}[GAN]{generative adversarial network}
\acrodef{CGAN}[CGAN]{conditional generative adversarial network}
\acrodef{ACGAN}[ACGAN]{auxilliary conditional generative adversarial network}
\acrodef{DCGAN}[DCGAN]{deep convolutional generative adversarial network}
\acrodef{CNN}[CNN]{convolutional neural networks}
\acrodef{BBH}[BBH]{binary black hole}

@@

% introduce \ac{GW} astrophsyics
%
\ac{GW} astronomy is now an established field, starting with the first
detection of a binary black hole merger \cite{Abbott2016} in September 2015.
Following this, the first and second observations runs (O1 and O2) of Advanced
LIGO and Advanced Virgo \cite{Prospects-dets, AdvLIGO, AdvLIGO2, AdvVIRGO} ~\chris{need refs for the detectors, see here
https://pnp.ligo.org/ppcomm/standard\_references.html} ~\jordan{\ding{51}} reported several more
mergers \cite{Abbott2016a, Abbott2017, Abbott2017a, Abbott2017b}. On August
2017 a binary neutron star merger was observed alongside its electromagnetic
counterpart for the first time, giving rise to multimessenger gravitational
wave astronomy.

% introduce burst searches
%
GW bursts are transient signals of typically short duration ($<$ 1s) whose
waveforms are not accurately modelled or are complex to re-produce.
Astrophysical sources for such transients include: Core collapse supernova,
Neutron star instabilities, Fallback accretion onto a neutron star, Non
axisymetric deformation in magnetars, Pulsar glitches and Neutron star
post-mergers. Currently, \ac{GW}s are detected using matched filtering, \cite{Owen1998} a process which requires template banks of possible GW signals. As \ac{GW} bursts are un-modelled there are no templates available and so these signals are undetectable through this scheme. ~\chris{OK, but logically dodgy. You make it
sound like we could use a template scheme but it would be insensitive. The
point is that we don't have templates in the first place.} \jordan{changed this but still reads clunky} Instead, detection involves distinguishing
the signal from detector noise. This is only possible if the detector noise is
well characterised and the candidate signal can be differentiated from
systematic or environmental glitches. As such, \ac{GW} burst searches rely on
an astrophysical burst signature appearing in multiple detectors at similar times. ~\chris{at
similar times?}. \ding{51}

% Discuss the family of burst waveforms currently used and why - not in detail, just
% an introduction 
%
\ac{GW} burst algorithms \cite{Klimenko_2008, Aso_2008}~\chris{many? needs
more than 2 refs for many.} are tested and tuned using model waveforms that may
or may not have astrophysical significance but have easy to define parameters
and share characteristics of real bursts that is enough to simulate a \ac{GW} passing ~\chris{fancy sounding but I'm not sure what
non-stationary deviations means.} between detectors. 
\begin{comment}Such waveforms may have
long-duration, short bandwidth (ringdowns), long-duration, large bandwidth
(inspirals) and many algorithms make use of sine-Gaussians: a Gaussian
modulated sine wave that is characterised by it's central frequency and narrow
bandwidth.~\chris{not the time to try to describe the types of burst waveforms.
Also, be careful with satying things like ringdowns have narrow bandwidth. I
know we specify a single frequency but becuase of the short duration, the
signal is broad band. Look at the FFT} This makes it a great tool for
diagnosing LIGOs sensitivity to frequency.~\chris{strange unfinished sentence.} 
\end{comment}

% Introduce ML techniques in GWs
%
\chris{Introduce ML techniques in GWs}

With the expectation that there will be many more \ac{GW} detections in the future, researches are starting to realise the need for fast and efficient \ac{GW} analysis to compete with the rising number of detections. Machine learning has shown great potential for \ac{GW} analysis in areas of detection \cite{george2016deep,Gabbard2017,Gebhard_2019,Krastev_2020}, glitch classification \cite{Bahaadini, George_2018,Razzano_2018} and parameter estimation \cite{gabbard2019bayesian,shen2019deterministic,green2020gravitationalwave}. Applying these machine learning algorithms show similar results to matched filtering techniques and have the benefit of being orders of magnitude faster.

% What this paper does on GANs in 1 paragraph
%
\chris{What this paper does on GANs in 1 paragraph}
In this work we aim to explore the use of machine learning to generate and interpret
unmodelled \ac{GW} burst waveforms. \jordan{i haven't mentioned gans yet do i explain it here briefly?}
% the point of the paper
%
\chris{describe the structure of the paper}
This paper is organised as follows: \jordan{unstructured at the moment}

We aim to explore the use of machine learning in generating and interpreting
these mock \ac{GW} burst signals. Neural networks have shown to replicate the
sensitivities of matched filtering in \ac{GW} detection \cite{Gabbard2017} and rapid
parameter estimation \cite{gabbard2019bayesian}, however these methods have
primarily been focused on binary black hole signals and have not yet expanded
to burst examples.~\chris{OK, thanks for the referdndes but you should probably
have more refs to other ML GW papers. I think you should dedicate at least one
paragrpah to GW ML efforts. The final paragrapgh of the section should
introduce the paper structure. The penultimate paragraph should expain what the
point of this paper is - in non technical terms. This is all outlined in the
bullet points (hidden) at the start of this section.}

@@

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, Machine Learning (ML) aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer Figure \ref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

@@

% How a GAN works
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
that is set up to distinguish between real and fake data and a generator that
produces synthetic reproductions of the real data. The generator performs a
mapping from an input noise vector $\mathbf{z}$, that is usually sampled from a n-dimensional Gaussian space known as the latent space, to its representation of the
data and the discriminator maps its input $\mathbf{x}$ to a probability that
the input came form either the training data or generator.  During training,
the discriminator is given a batch of samples that contains one half real data
and one half fake data which it then makes predictions on. The loss for the
discriminator is calculated by comparing its predictions to the labelled data
through the binary cross-entropy function. \jordan{note to myself: This is important since eqn 2 is derived from it (thats where the logs come from)} The training process of a \ac{GAN}
alternatively updates the weights~\chris{too much knowledge assumed - what are
weights, they haven't been defined} \jordan{defined above but went on a bit, should i go into back prop and optimizers?} of the discriminator and generator based on
information on its competitors loss function. This loss of discriminator is
used to update the weights of the generator to produce more realistic samples
of the input distribution, the loss of the generator encourages the
discriminator to update its classification abilities. GANs seek to find an equilibrium between the generator and discriminator in the form of a two-player game that can be summarised by the following:

@@

where, V is an objective function that is to be optimized. Objective functions are a general term for either a loss function (to be minimized) or a negative loss function (to be maximised). D is the discriminator, G is the generator and ${\mathbf{x} \sim p_{\text{r}}(\mathbf{x})}$, ${\mathbf{z} \sim p_{\text{z}}(\mathbf{z})}$ are samples taken from the training data and generated data respectfully. In practice this approach was found to be unstable. Early on in training the generators weights are initialised randomly meaning that the early generations are noisy and uninformative. This means that the discriminator can easily learn to spot the fake generations and so the discriminator wins and the generator cannot continue to improve. To overcome this the framing of the generator is often changed. Rather than minimizing $\text{log}(1-D(G(\mathbf{z})))$ the objective now is to maximise $\text{log}(D(G(\mathbf{z})))$, that is, to maximise the probability of the generations being predicted as real. The idea being that we always want the loosing side to be able to leverage from its competitor. 
%
~\chris{In line with explanding the explanation a bit more, really take the time
to state what controls the functions and also deine ALL mathematical variables.
What are $D, G, V\ldots$ Why are you taking the logs? What situation so you
start with? What do you end up with? This is a nice start but beef it up.}

@@

%
\begin{comment}adds structure to the latent space by providing the
generator with a class or attribute label. The generator learns to segment the
latent space by clustering distributions which have similar properties, making
a point in latent space conditional on a class~\chris{are we sure about this?
all classes share the same latent space and I don't think that the space gets
segmented into areas corresponding to different classes - I do think that it
gets broken down into spaces with different intrinsic physical properties e.g.,
high frequency, long duration, etc...}.
\end{comment}
%
This idea was extended further with
\acp{ACGAN}~\cite{odena2016conditional} that require the discriminator to
output a probability of data belonging to each class. A pictorial
representation on the differences between these approaches is shown in
Fig.~\ref{fig:gan_comparison} \jordan{self note: Probably move away from ACGAN part, standalone aux classifiers can be trained later if that's what you want}.

\begin{figure}
    \centering
        \input{figures/GAN_diagram_vertical}%     without .tex extension
        \input{figures/ACGAN_diagram_vertical}%     without .tex extension
    \caption{Comparison of the original GAN method and the Auxiliary
Conditional-GAN method. For ACGANs the training data requires a label denoting
its class that is also fed to the generator which then learns to generate
waveforms based on the input label. Additionaly, the discrminator learns to
classify which class the signal belongs to.} \label{fig:gan_comparison}
\end{figure}

\chris{this section is supposed to be on ACGANs but it only mentions them in
the final sentence. Again, like the previous section, a bit more explanation
would be good. Does ths maths change? Could you have a new version of Eq 1 for
CGANs and ACGANs? Also, if would be good to have a statement that you will be
using one of these methods in this paper.}

@@

% introduce the training data
%
\ac{GW} burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveform reconstruction reliant on coincident
signals within multiple detectors~\chris{burst detection algorithms do not need
waveform reconstruction for searches - it is an add on feature.}. We propose a
signal generation scheme utilizing~\chris{utilizing certainly is fancier than
using. Dont go overboard with the language.} \acp{GAN} trained on burst-like
waveforms. The \ac{GAN}~\chris{is it a GAN or is it a CGAN or an ACGAN? Be
clearer} is trained on five signal morphology's spanning a range of prior
parameters. The families are:~\chris{I know they don't all share the same
parameters but it might be sensible to make a table of the prior ranges rather
than haviong it all stated in text.}
%
\begin{itemize}
%
\item Sine-Gaussian:\\
\chris{don't just start things with an equation}		
%
\begin{equation}
\label{eqn:sg}
h(t) = A \exp\bigg[ - \frac{(t-t_{0})^2}{\tau^2} \bigg] \sin (2 \pi f_0 (t-t_0))
\end{equation}	
%		
where $f_0$ is the central frequency that takes values between 30 Hz - 50 Hz.
$\tau$ is the decay parameter chosen to be between 60s$^{-1}$  and 15s$^{-1}$
and the starting epoch chosen between 0.2s - 0.8s. Sine-Gaussian waveforms have
a similar form to those produced by the merger of relatively high mass binary
black holes for which the inspiral phase occurs at frequencies below
ground-based detector sensitivities.
%
\item Ring-down:\\
\chris{don't just start things with an equation}
%
\begin{equation}
h(t) = A \exp \bigg[-\frac{(t-t_0)}{\tau}\bigg]\sin(2 \pi f_0 (t-t_0))
\end{equation}
%		
For ring-down signals the parameters for $f_0, \tau, t_0$ are chosen between;
30 Hz-50Hz, 0.02-0.1 and 0.1s-0.8s respectfully. These signals aim to replicate
the post-merger late stages of binary coalescence.
%
\item White-noise bursts:\\
\chris{where is the waveform equation?}
These signals are produced by inserting samples of random Gaussian noise with
zero mean and a variance of 0.1 at a random point between 0.2s - 0.8s. The
source of this signal type can be attributed to core-collapse
supernova~\chris{is this correct? these waveofms are approximations so best not
to say that they correspond directly to physical phenomena}.
%
\item Gaussian pulse:\\
\chris{where is the waveform equation?}
A simple Gasssian pulse waveform with $t_0$ ranging from 0.2s to 0.8s and
$\tau$ between  100$^{-1}$ and 20$^{-1}$

\item Binary black holes (BBH): \\
\ac{BBH} signals are simulated using IMRPhenomD waveform routine from
LALSuite~\chris{references for both the waveofrm and lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70]M$_0$~\chris{not $M_0$ but $M_{\odot}$} with zero spins
and we fix m$_1$ $>$ m$_2$. The mass distribution is approximated by a power
law with index of 1.6 https://arxiv.org/abs/1811.12940~\chris{correct;ly
reference this}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. The peaks of the
waveforms are set to be within [0.75,0.95]s of the 1s time interval.~\chris{for
all the signal models please check that the prior ranges and distributions are
consistent with the presented results.} 
%
\end{itemize}

\chris{you seem to be missing a description of the antenna patterns and time
delay that you apply to each waveform. It should go here. You should also
include all other technical information like how you normalise the amplitudes
and how you select random sky positions, how the data is represented to the
network (as an image or as channels) how many detectors do you use? etc... }

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/train_gen.pdf}
    \caption{Examples of simulated \ac{GW} burst signals. Top row shows
examples from the training set. From left to right: Sine-Gaussian, Ringdown,
White-noise burst, Gaussian pulse, Binary black hole merger. The bottom row
shows the conditional generations from the GAN.~\chris{slightly more
explanation. The reader may be concerned that the top and bottom do not look
the same. Should they? They may wonder why is there only one waveform and not
one per detector?} } \label{fig:train} \end{figure}

@@

% Introduce some architecture features
%
Extensions to the original \ac{GAN} method such as the
\ac{DCGAN}~\cite{Radford2015} have been widely praised for
building~\chris{enabling?} a stable \ac{GAN} architecture. Modern \ac{GAN}
research~\chris{isn't all GAN research modern?} favours a fully convolution
neural network which replace upsampling procedures like maxpooling to strided
convolutions~\chris{a couple of things here. Max pooling and striding hasn't
been introduced, and why do max-pooling and striding upsample? I thought they
downsampled, don't they? }. \acp{CNN} are designed to work with grid-like
structures that exhibit strong local spatial dependencies~\chris{great but that
sounds like a computer science paper. Make it easier for the physicists to
understand.}. Although most work with \acp{CNN} involve image based data, they
can be applied to other spatially adjacent data types such as time-series and
text items. Audio synthesis work~\cite{DBLP:journals/corr/abs-1809-11096} has
showed that \acp{GAN} can achieve recognisable audio within a few
hours~\chris{ambiguous. A few hours of training?}.

% specifically define our architecture
%
We adopt the suggestions of these papers~\chris{which papers?}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator has two output layers: the
first output is a single node activated by a Sigmoid that can be interpreted as
the realness of the the signal, the second output is 5 nodes activated using
the softmax function predicting the class of the input.  This model is trained
with binary cross entropy for the first output and sparse categorical
cross-entropy for the second output.~\chris{OK, very technically tight but too
computer sciencey. Try to make it human readbale for a physicist and refer to
the table of parameters in the appendix.}

@@

% Explain the effect of the sky position 
%
\chris{Why is this not described in the section defining the signal waveforms?
There may be a good reason so it may well fit best here. I'm not sure} We
consider a two detector case~\chris{why 2 detector and not 3?} in which the
generator is trained to output two identical signals with a physical time shift
representing the time of flight between detectors and the
suppression~\chris{not a supression. It's just the antenna response.} of
amplitudes due to interaction with the detector. To achieve this, the generator
is trained to output a single waveform that is then put through a non-trainable
``response" layer before the output of the generator. This layer creates a copy
of its input, shifts it along in time and applies antenna responses to both the
input and shifted signal. The time shift is achieved by converting the waveform
to the frequency domain, multiplying by $e^{2 \pi i f \Delta t}$, where $\Delta
t$ is the time shift and converting~\chris{converting? what do you mean?} back
to the time domain. After which, both original and shifted signals are
multiplied by their respective  antenna responses. Both $\Delta t$ and the
antenna responses are calculated using the LALSimulation~\cite{lalsuite}
package and we choose Hanford and Livingston as detection points~\chris{just
say the detectors}. The result is a $2\times 1024$ times series waveform
representing an overlay data stream~\chris{what the hell is an overlay data
stream?} from two detector outputs. This scheme is also used in generating the
training set.~\chris{OK, first of all you should define some equations here
rather than just saying everything in words. Why is it here that you reveal the
data sampling rate? Surely that is an important quantity that shouldn't just be
stated in this subsection. What about the random sky position that is input to
the "box"? How is the sky sampled? What about the wrapping from the tim shift?
Don't you apply a window? The "box" or "response" layer should be visible in
the network diagram.}

>>>>

\acrodef{GW}[GW]{Gravitational wave}
\acrodef{GAN}[GAN]{generative adversarial network}
\acrodef{CGAN}[CGAN]{conditional generative adversarial network}
\acrodef{ACGAN}[ACGAN]{auxilliary conditional generative adversarial network}
\acrodef{DCGAN}[DCGAN]{deep convolutional generative adversarial network}
\acrodef{CNN}[CNN]{convolutional neural networks}
\acrodef{BBH}[BBH]{binary black hole}

@@

% introduce \ac{GW} astrophsyics
%
Gravitational wave astronomy is now an established field, starting with the first
detection of a binary black hole merger \cite{Abbott2016} in September 2015.
Following this, the first and second observations runs (O1 and O2) of Advanced
LIGO and Advanced Virgo \cite{Prospects-dets, AdvLIGO, AdvLIGO2, AdvVIRGO} reported several more Compact Binary Coalescence (CBC)
mergers \cite{Abbott2016a, Abbott2017, Abbott2017a, Abbott2017b}. On August
2017 a binary neutron star merger was observed alongside its electromagnetic
counterpart for the first time, giving rise to multimessenger gravitational
wave astronomy.

% introduce burst searches
%
With these successes and continued upgrades to the detectors, further detections of CBCs are expected to be commonplace in a few short years. Another group of GW signals that has thus far been undetectable is GW ``bursts".  GW bursts are transient signals of typically short duration ($<$ 1s) whose
waveforms are not accurately modelled or are complex to re-produce.
Astrophysical sources for such transients include: Core collapse supernova,
Neutron star instabilities, Fallback accretion onto a neutron star, Non
axisymetric deformation in magnetars, Pulsar glitches and Neutron star
post-mergers.

GW detections so far have used a process called matched filtering, \cite{Owen1998}, where a large template bank of possible GW waveforms are compared to the detector outputs. In order to increase the chances of detection these template banks must span a large multi dimensional parameter space and the templates must be accurate, requiring significant computational cost.  GW bursts are un-modelled; therefore there are no templates available and so these signals are undetectable through this scheme. Instead, detection involves distinguishing
the signal from detector noise. Burst searches look for excess power contained in the time-frequency domain and rely on the astrophysical burst waveform appearing in multiple detectors at similar times. This is only possible if the detector noise is
well characterised and the candidate signal can be differentiated from
systematic or environmental glitches.

% Discuss the family of burst waveforms currently used and why - not in detail, just
% an introduction 
%
\ac{GW} burst algorithms \cite{Klimenko_2008, Aso_2008} are tested and tuned using model waveforms that may
or may not have astrophysical significance but have easy to define parameters
and share characteristics of real bursts that is enough to simulate a \ac{GW} passing between detectors. Such waveforms include sine-Gaussians: a Gaussian modulated sine wave that is by it's central frequency and decay parameter. Bandlimited white-noise bursts: white noise that is contained within a certain frequency range and ring-downs which mimic the damped oscillations after a CBC merger.

\begin{comment}Such waveforms may have
long-duration, short bandwidth (ringdowns), long-duration, large bandwidth
(inspirals) and many algorithms make use of sine-Gaussians: a Gaussian
modulated sine wave that is characterised by it's central frequency and narrow
bandwidth.~\chris{not the time to try to describe the types of burst waveforms.
Also, be careful with satying things like ringdowns have narrow bandwidth. I
know we specify a single frequency but becuase of the short duration, the
signal is broad band. Look at the FFT} This makes it a great tool for
diagnosing LIGOs sensitivity to frequency.~\chris{strange unfinished sentence.} 
\end{comment}

% Introduce ML techniques in GWs
%

With the expectation that there will be many more \ac{GW} detections in the future, researches are starting to realise the need for fast and efficient \ac{GW} analysis to compete with the rising number of detections. While still in its infancy, Machine Learning (ML) with GW has already shown great potential for overcoming data bottlenecks. \ac{GW} analysis in areas of detection \cite{george2016deep,Gabbard2017,Gebhard_2019,Krastev_2020}, glitch classification \cite{Bahaadini, George_2018,Razzano_2018} and parameter estimation \cite{gabbard2019bayesian,shen2019deterministic,green2020gravitationalwave} have shown the success of in ML pattern recognition. Applying these algorithms show similar results to matched filtering techniques and have the benefit of being orders of magnitude faster.

% What this paper does on GANs in 1 paragraph
%

In this work we aim to explore the use of machine learning to generate and interpret
unmodelled \ac{GW} burst waveforms. Using the generative machine learning model: Generative Adversarial Networks (GANs), we train on five classes of known burst morphologies in the time domain. Working on the assumption that GANs construct smooth n-dimension vector spaces between it's input and output, we can then explore the space between the five classes to construct new unmodelled waveforms. As all the computationally expensive processes occur during training, the learned model after training is able to produce waveforms much faster than and in fact produce waveforms that are impossible to produce with current techniques. These new varieties of waveforms can then be used to diagnose detection algorithms and gain new insight into sources of \ac{GW} bursts. 
% the point of the paper
%

This paper is organised as follows: \jordan{unstructured at the moment}

@@

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, ML aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer Figure \ref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

@@

% How a GAN works
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
that is set up to distinguish between real and fake data and a generator that
produces synthetic reproductions of the real data. The generator performs a
mapping from an input noise vector $\mathbf{z}$, that is usually sampled from a n-dimensional Gaussian space known as the latent space, to its representation of the
data and the discriminator maps its input $\mathbf{x}$ to a probability that
the input came form either the training data or generator.  During training,
the discriminator is given a batch of samples that contains one half real data
and one half fake data which it then makes predictions on. The loss for the
discriminator is calculated by comparing its predictions to the labelled data
through the binary cross-entropy function. \jordan{note to myself: This is important since eqn 2 is derived from it (thats where the logs come from)} The training process of a \ac{GAN}
alternatively updates the weights of the discriminator and generator based on
information on its competitors loss function. This loss of discriminator is
used to update the weights of the generator to produce more realistic samples
of the input distribution, the loss of the generator encourages the
discriminator to update its classification abilities. GANs seek to find an equilibrium between the generator and discriminator in the form of a two-player game that can be summarised by the following:

@@

where, V is an objective function that is to be optimized. Objective functions are a general term for either a loss function (to be minimized) or a negative loss function (to be maximised). D is the discriminator, G is the generator and ${\mathbf{x} \sim p_{\text{r}}(\mathbf{x})}$, ${\mathbf{z} \sim p_{\text{z}}(\mathbf{z})}$ are samples taken from the training data and generated data respectfully. In practice this approach was found to be unstable. Early on in training the generators weights are initialised randomly meaning that the early generations are noisy and uninformative. This means that the discriminator can easily learn to spot the fake generations and so the discriminator wins and the generator cannot continue to improve. To overcome this the framing of the generator is often changed. Rather than minimizing $\text{log}(1-D(G(\mathbf{z})))$ the objective now is to maximise $\text{log}(D(G(\mathbf{z})))$, that is, to maximise the probability of the generations being predicted as real. The idea being that we always want the loosing side to be able to leverage from its competitor. 
%

@@

This simple addition has shown to work well in practice, for instance in image-to-image translation \cite{isola2016imagetoimage}. We will be using a conditional GAN for this study.
%
\begin{comment}adds structure to the latent space by providing the
generator with a class or attribute label. The generator learns to segment the
latent space by clustering distributions which have similar properties, making
a point in latent space conditional on a class~\chris{are we sure about this?
all classes share the same latent space and I don't think that the space gets
segmented into areas corresponding to different classes - I do think that it
gets broken down into spaces with different intrinsic physical properties e.g.,
high frequency, long duration, etc...}.
\end{comment}
%
\begin{comment}
This idea was extended further with
\acp{ACGAN}~\cite{odena2016conditional} that require the discriminator to
output a probability of data belonging to each class. A pictorial
representation on the differences between these approaches is shown in
Fig.~\ref{fig:gan_comparison} \jordan{self note: Probably move away from ACGAN part, standalone aux classifiers can be trained later if that's what you want}. 
\end{comment}

\begin{figure}
    \centering
        \input{figures/GAN_diagram_vertical}%     without .tex extension
        \input{figures/ACGAN_diagram_vertical}%     without .tex extension
    \caption{Comparison of the original GAN method and the Auxiliary
Conditional-GAN method. For ACGANs the training data requires a label denoting
its class that is also fed to the generator which then learns to generate
waveforms based on the input label. Additionaly, the discrminator learns to
classify which class the signal belongs to. \jordan{change to cGAN}} \label{fig:gan_comparison}
\end{figure}

@@

% introduce the training data
%

\begin{table}[hb]
\centering
\caption{Burst training parameters}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($M_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A \\
White-noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
Gaussian pulse & N/A & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
BBH & N/A & N/A & N/A & 5 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

GW burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveforms signals appearing within multiple detectors We propose a
signal generation scheme using \acp{GAN} trained on burst-like
waveforms. BurstGAN is a conditional GAN that is trained on five signal morphology's spanning a range of prior
parameters. The parameter ranges can be seen in Table \ref{Tab:training_parms} The families are: sine-Gaussian $h(t) = A \exp\bigg[ - \frac{(t-t_{0})^2}{\tau^2} \bigg] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian envelop characterised by a central frequency, $f_0$; Ring-down, $h(t) = A \exp \bigg[-\frac{(t-t_0)}{\tau}\bigg]\sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$, white-noise bursts, with bounded frequency bandwidth $\Delta f$ and duration $\tau$, gaussian pulse $h(t) = \exp(-\frac{t^2}{\tau^2})$ with duration $\tau$ and binary-black hole inspirals which are simulated using IMRPhenomD waveform \cite{Khan_2016} routine from
LALSuite \cite{lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70]$M_{\odot}$ with zero spins
and we fix m$_1$ $>$ m$_2$. The mass distribution is approximated by a power
law with index of 1.6 \cite{Abbott_2019}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. The peaks of the
waveforms are set to be within [0.4,0.6]s of the 1s time interval.

All training waveforms are sampled at 1024 Hz.
%

% Explain the effect of the sky position 
%
We consider a two detector case~\chris{why 2 detector and not 3?} in which the
generator is trained to output two identical signals with a physical time shift
representing the time of flight between detectors H1 and L1 To achieve this, the generator
is trained to output a single waveform that is then put through a non-trainable
``response" layer before the output of the generator. This layer takes as input, the generated waveform, time delay and antenna responses. The sky location is chosen by drawing from a uniform distribution across the whole sky and the time delay, and antenna responses generated by LALSimulation~\cite{lalsuite}.

This response layer Fourier transforms the waveform
to the frequency domain, multiplies this by $e^{2 \pi i f \Delta t}$, where $\Delta
t$ is the time shift and transforming back
to the time domain. After which, both original and shifted signals are
multiplied by their respective  antenna responses. The output is two 1024 Hz times series waveform split into two channels.
This scheme is also used in generating the
training set.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/train_gen.pdf}
    \caption{Examples of simulated \ac{GW} burst signals. Top row shows
examples from the training set. From left to right: Sine-Gaussian, Ringdown,
White-noise burst, Gaussian pulse, Binary black hole merger. The bottom row
shows the conditional generations from the GAN. Each plots contains only one detector output for ease of viewing.}
\label{fig:train} \end{figure}

@@

% Introduce some architecture features
%
Extensions to the original \ac{GAN} method such as the
\ac{DCGAN}~\cite{Radford2015} have been widely praised for
enabling a stable \ac{GAN} architecture. Most GANs now replace fully connected layers, where each neuron is connect to all the neurons in the previous and next layer, with convolutional layers. \ac{CNN} are designed to work with grid-like structures with close local dependencies like image and text based data, however, there are examples of audio synthesis work \cite{DBLP:journals/corr/abs-1809-11096}.

% specifically define our architecture
%
We adopt the suggestions of \cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator has two output layers: the
first output is a single node activated by a Sigmoid that can be interpreted as
the realness of the the signal, the second output is 5 nodes activated using
the softmax function predicting the class of the input.  This model is trained
with binary cross entropy for the first output and sparse categorical
cross-entropy for the second output. The full architecture description can be seen in Table \ref{Tab:hyperparameters}.~\chris{OK, very technically tight but too
computer sciencey. Try to make it human readbale for a physicist and refer to
the table of parameters in the appendix.}

@@
